Spatial Reasoning and Statistical Graphics

### 2014-10-17 Paper/Draft1/TestingVisualAptitude.Rnw
We don't know if the lineups are null or not...
scaling

Between cognitive tasks, participants will also complete three blocks of 20 lineups each. These lineups have been previously tested \citep{hofmann2012graphical} and include some null lineups (i.e. lineups without a target plot). Participants %will 
have 5 minutes to complete each block of 20 lineups. Figure \ref{fig:lineup} shows a sample lineup of box plots.

@@

\newtext{The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995manual} and comparison data are available from the validation of these factors. The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).}

% \comment{If test scores have the same ranges, we could include a mini table of means and standard deviations of the four tests, and discuss whether these findings are consistent to how people usually score (except for the females we have about the same population).}
\begin{savenotes}
\begin{table}[ht]
\centering
\begin{tabular}{rlll}
\hline
  & Card Rotation & Paper Folding & Figure Classification \\\hline
ISU Students & \Sexpr{round(mean(ans.summary.unscaled$card_rot), 2)} (\Sexpr{round(sd(ans.summary.unscaled$card_rot), 1)}) 
             & \Sexpr{round(mean(ans.summary.unscaled$folding), 2)} (\Sexpr{round(sd(ans.summary.unscaled$folding), 1)})
             & \Sexpr{round(mean(ans.summary.unscaled$fig_class), 2)} (\Sexpr{round(sd(ans.summary.unscaled$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}\\
Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})
              & 13.8 (4.5)
              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 2)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617/2), 1)})\footnote{Averages calculated assuming 294 males and 323 females.}\\
Unscaled Scores & 44.0 (24.6)\footnote{Data from Part I only.}
                & 13.8 (4.5)
                & M:~120.0~(30.0), F:~114.9~(27.8)\\
{\footnotesize Population}    
              & \parbox[t]{.2\textwidth}{\footnotesize $\approx$ 550 male Naval recruits} 
              & \parbox[t]{.2\textwidth}{\footnotesize 46 college students\\(1963~version)}
              & \parbox[t]{.3\textwidth}{\footnotesize Suburban 11-12th grade students\\(288-300 males, 317-329 females)}\\\hline
\end{tabular}

@@

\newtext{Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations. Once values have been adjusted to accommodate different test procedures (some data is reported for a single part of a two-part test, for instance), it is evident that Iowa State undergraduates scored at about the same level as other similar demographics. In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.}\footnote{That is, if comparison data was available for 9th grade students and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, as they are closer in age to college students. When data was available from college students and Army enlistees, we have compared ISU students to other college students, as college students are more likely to have similar gender distribution to ISU students.}

\newtext{As we have established that the results obtained for the ETS tests are similar to other studies, we will now compare the results to the lineups also tested in this study. To facilitate this goal, for the remainder of this analysis, we will scale the test results so that the ranges and units of test scores are comparable.} Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
This leads to a theoretical range of $X_{n,k}$ of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of

>>>>

Between cognitive tasks, participants will also complete three blocks of 20 lineups each. These lineups have been previously tested \citep{hofmann2012graphical} %and include some null lineups (i.e. lineups without a target plot). 
. Participants %will 
have 5 minutes to complete each block of 20 lineups. Figure \ref{fig:lineup} shows a sample lineup of box plots.

@@

The advantage of using tests from the Kit of Factor Referenced Cognitive tests \citep{ekstrom1976manual} is that the tests are extremely well studied \citep{voyer1995manual}
\comment{the bib file is not updated on the repo}
and comparison data are available from the validation of these factors. The card rotation, paper folding, and figure classification tests have been validated using different populations, many of which are demographically similar to Iowa State students (naval recruits, college students, late high-school students, and 9th grade students).

% \comment{If test scores have the same ranges, we could include a mini table of means and standard deviations of the four tests, and discuss whether these findings are consistent to how people usually score (except for the females we have about the same population).}
\begin{savenotes}
\begin{table}[ht]
\centering
\begin{tabular}{rlll}
\hline
  & Card Rotation & Paper Folding & Figure Classification \\\hline
ISU Students & \Sexpr{round(mean(ans.summary.unscaled$card_rot), 1)} (\Sexpr{round(sd(ans.summary.unscaled$card_rot), 1)}) 
             & \Sexpr{round(mean(ans.summary.unscaled$folding), 1)} (\Sexpr{round(sd(ans.summary.unscaled$folding), 1)})
             & \Sexpr{round(mean(ans.summary.unscaled$fig_class), 1)} (\Sexpr{round(sd(ans.summary.unscaled$fig_class), 1)})\footnote{ISU students took only Part I due to time constraints.}\\
Scaled Scores & 88.0 (\Sexpr{round(sqrt((24.6^2)*2), 1)})
              & 13.8 (4.5)
              & \Sexpr{round((120.0*294 + 114.9*323)/617/2, 1)} (\Sexpr{round(sqrt((30^2*294 + 27.8^2*323)/617)/2, 1)})\footnote{Averages calculated assuming 294 males and 323 females.}\\
Unscaled Scores & 44.0 (24.6)\footnote{Data from Part I only.}
                & 13.8 (4.5)
                & M:~120.0~(30.0), F:~114.9~(27.8)\\
{\footnotesize Population}    
              & \parbox[t]{.2\textwidth}{\footnotesize approx, 550 male \\
Naval recruits} 
              & \parbox[t]{.2\textwidth}{\footnotesize 46 college students\\(1963~version)}
              & \parbox[t]{.3\textwidth}{\footnotesize Suburban 11-12th grade students\\(288-300 males, 317-329 females)}\\\hline
\end{tabular}

@@

Table \ref{tab:scorecomparison} shows mean scores and standard deviation for ISU students and other populations. Once values have been adjusted to accommodate different test procedures (some data is reported for a single part of a two-part test, for instance), it is evident that Iowa State undergraduates scored at about the same level as other similar demographics. In fact, both means and standard deviations of ISU students' scores are similar to the comparison groups, which were chosen from available demographic groups based on population similarity.\footnote{If comparison data was available for 9th and 12th grade students, we have compared Iowa State students' scores with the 12th grade students, to closer in age to college students. When data was available from college students and Army enlistees, we have compared ISU students to other college students, as college students are more likely to have similar gender distribution to ISU students.}
\comment{Could you include some more details on how to get from the Unscaled scores to the scaled versions? I'm assuming that for the Card Rotation, the two parts have equal size, explaining the doubled score, but why not double the standard deviation? $\text{Var}(aX) = a^2 \text{Var}(X)$      I just don't quite follow the code for the combined scores ... it would be worthwhile to include a bit more detail and maybe put that in an appendix to the paper. If you want to, you could instead extend the footnotes a bit for now. ... so just let me try this: Let's assume that $X_M$ and $X_F$ are the scores based on 294 men, and 323 women (are those midscore s of the intervals reported?). with corresponding $s_M$ and $s_F$ for the reported standard deviation of the full test. Then half of the test would have half of the scores with half of the standard deviation. we have a breakdown of 17 girls and 19 boys. $
X = (X_M/2 \cdot 294 + X_F/2 \cdot 323) /(294+323) = \Sexpr{round(120.0/2*294/617 + 114.9/2*323/617, 1)}
$}
As we have established that the results obtained for the ETS tests are similar to other studies, we will now compare the results to the lineups also tested in this study. To facilitate this goal, for the remainder of this analysis, we will scale the test results so that the ranges and units of test scores are comparable.
Let $X_{n, k}$ be a participant's score on a test consisting of $n$ questions with $k$ answers each, out of which only one is correct.
This leads to a theoretical range of $X_{n,k}$ of $[-n/(k-1), n]$ and, under an additional assumption of random guessing, a variance of
